---
title: "Exploratory Data Analysis - Fed Rate Prediction"
output: html_document
---
```{r}
install.packages("corrplot")
install.packages("GGally")
```
# --------------------------------
# 1️⃣ Load Libraries
# --------------------------------
```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(lubridate)
library(corrplot)
library(GGally)
```

# --------------------------------
# 2️⃣ Load Clean Dataset
# --------------------------------
```{r}
data_dir <- "../data/"
fred_data <- read_csv(file.path(data_dir, "fred_clean.csv"))

glimpse(fred_data)
```

# --------------------------------
# 📊 3️⃣ Summary Statistics
# --------------------------------
```{r}
summary(fred_data)
fred_data %>%
  select(where(is.numeric)) %>%
  summarise(across(everything(), list(
    mean = mean,
    sd = sd,
    min = min,
    max = max
  ), na.rm = TRUE))
```

# --------------------------------
# 📈 4️⃣ Time Series Visualization
# --------------------------------
*Federal Funds Rate Over Time*
```{r}
ggplot(fred_data, aes(x = date, y = FEDFUNDS)) +
  geom_line(color = "blue") +
  labs(title = "Federal Funds Rate Over Time",
       x = "Date",
       y = "FEDFUNDS") +
  theme_minimal()
```
*Inflation (CPI) Over Time*
```{r}
ggplot(fred_data, aes(x = date, y = CPIAUCSL)) +
  geom_line(color = "red") +
  labs(title = "CPI Over Time",
       x = "Date",
       y = "CPI") +
  theme_minimal()
```
# --------------------------------
# 📉 5️⃣ Yield Spread Analysis
# --------------------------------
```{r}
ggplot(fred_data, aes(x = date, y = YieldSpread)) +
  geom_line(color = "darkgreen") +
  labs(title = "10Y - 2Y Yield Spread",
       x = "Date",
       y = "Yield Spread") +
  theme_minimal()
```
# -----------------------------------------------
# 🔎 6️⃣ Correlation Matrix (Numerical Variables)
# -----------------------------------------------
```{r}
numeric_data <- fred_data %>%
  select(where(is.numeric)) %>%
  select(-FedRateChange)

cor_matrix <- cor(numeric_data, use = "complete.obs")

corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.6)
```
# --------------------------------------------------------
# 📊 7️⃣ Fed Rate Direction Distribution (Categorical)
# --------------------------------------------------------
```{r}
ggplot(fred_data, aes(x = RateDirection)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Distribution of Rate Direction",
       x = "Rate Direction",
       y = "Count") +
  theme_minimal()
```

# --------------------------------
# 📌 9️⃣ Pairwise Plot (Important)
# --------------------------------
```{r}
selected_vars <- fred_data %>%
  select(FEDFUNDS, CPIAUCSL, UNRATE, YieldSpread, VIXCLS)

GGally::ggpairs(selected_vars)
```

## Key Observations

- FEDFUNDS shows clear cyclical behavior.
- YieldSpread becomes negative before recession periods.
- CPI shows upward long-term trend.
- Certain variables show high correlation, indicating possible multicollinearity.
